Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Filling in more gaps
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@15 db7c848b-7755-4df2-b763-8796797afa52

application, renderer, view data

@@

The initial version of Equalizer implemented sort-first (2D), sort-last (DB) and
stereo (EYE) decomposition. In the following we present new decomposition modes
and motivate their use case.

@@

GPU-CPU transfer plugins with optional compression (YUV, RLE) linked to CPU
compression for network transfer \cite{MEP:10}

@@

Per-node sync, draw sync, async

>>>>

Sequal is a simplification layer for Equalizer. It is based on the realization
that while fully expressive, the verbatim abstraction layer of nodes, pipes,
windows and channels in Equalizer requires significant learning to fully
understand and exploit. In reality, a higher abstraction of \textsf{Application}
and \textsf{Renderer} is sufficient for most use cases. In Sequel, the
application class drives the configuration, and one renderer instance is created
for each (pipe) render thread. They also provide the natural place to store and
distribute data. Last, but not least, view data provides a convenient way to
manage multiple views by storing the camera, model or any other view-specific
information.

@@

The initial version of Equalizer implemented sort-first (2D), sort-last (DB) and
stereo (EYE) decomposition \cite{EMP:09}. In the following we present new
decomposition modes and motivate their use case.

@@

Compression for the compositing step in parallel rendering is critical for
performance. This not only applies to the well-researched network transfer step,
but also for the transfer between GPU and CPU. Equalizer supports a variety of
compression algorithms, from very fast RLE encoding, JPEG compression to YUV
subsampling on the GPU. These algorithms are implemented as runtime-loaded
plugins, allowing easy extension and customization to application-specific
compression. In \cite{MEP:10}, we show this to be a critical step for
interactive performance at scale.

@@

Different applications have different degrees on how decoupled and thread-safe
the rendering code is from the application logic. For full decoupling all
mutable data has to have a copy in each render thread, which is not feasible in
large data scenarios. To easily customize the synchronization of all threads on
a single process, Equalizer implements three threading modes:  Full
synchronization, draw synchronization and asynchronous. Note that the execution
between nodes is always asynchronous, for up to \textsf{latency} frames.

In full synchronization, all threads always execute the same frame, that is, the
render threads are unlocked after \textsf{Node::frameStart}, and the node is
blocked for all render threads to finish the frame before executing
\textsf{Node::frameFinish}. This allows the render threads to read shared data
from all their operations, but provides the slowest performance.

In draw synchronization, the node thread and all render threads are synchronized
for all \textsf{frameDraw} operations, that is, \textsf{Node::frameFinish} is
executed after the last channel is done drawing. This allows the render threads
to read shared data during their draw operation, but not during
compositing. Since compositing is typically independent of the rendered data,
this is the default mode. This mode allows to overlap compositing with data
synchronization on multi-GPU machines.

In asynchronous execution, all threads run asynchronously. Render threads may
work on different frames at any given time. This mode is the fastest, and
requires the application to have one instance of each mutable object in each
render thread. It is required for using time-multiplex compounds in
multithreaded rendering.
